---
title: "003_SPL_RaceRatio"
date: "2024-05-08"
output: html_document
---

Recodes race, calculates URM:White ratio by neighborhood and location_id, and maps the URM:White ratio

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
rm(list=ls())
library(tidyverse)
library(sf)
library(here)
```

# Import data
```{r}
SPL_1823 <- read.csv(here("data","SPL_1823.csv"))
SPL_1823_geo <- st_read(here("data", "SPL_1823_location.geojson")) # location_id shapefiles and lat/long coordinates
SEA_map <- st_read(here("data/raw","04_SPL_Seattle_Map.geojson")) # base SEA neighborhood map
```

# Recode the race column 
```{r}
SPL_1823 <- SPL_1823 %>%
  mutate(
    race=recode(staying_race_ethnicity,
                        "A" = "URM", # Asian
                        "B" = "URM", # Black or African American
                        "L" = "URM", # Latino/a or Hispanic
                        "M" = "URM", # Multiple, assumed to be multiracial
                        "N" = "URM", # Native American or Alaskan Native
                        "P" = "URM", # Pacific Island or Native Hawaiian
                        "U" = "Unsure", # Unsure
                        "W" = "White")) # White

table(SPL_1823$race)
table(SPL_1823$staying_race_ethnicity)

# Realized there were many blanks, so wanted to see where these were coming from

table(SPL_1823$staying_race_ethnicity[SPL_1823$study_id == "2023_Seattle_Citywide"])

# Confirmed that 2023 did not collect race data, so only have race data for 2018

SPL_18 <- SPL_1823 %>%
  filter(SPL_1823$study_id == "2018_Seattle_Citywide")

# Run tables again
table(SPL_18$race)
table(SPL_18$staying_race_ethnicity)

# Only 142 blanks now, which seems more normal

```

# Calculate URM/White ratio by neighborhood (S_HOOD)
```{r}
race_ratio <- SPL_18 %>%
  group_by(S_HOOD, race) %>%
  dplyr::summarise(count= n()) %>%
  pivot_wider(
    id_cols = S_HOOD,
    names_from = race,
    values_from = count,
    names_prefix = "Observed_"
  ) %>%
  mutate(race_ratio = round(Observed_URM/Observed_White, 2)) %>%
  select(S_HOOD, race_ratio)
```

# Calculate women/men ratio by observation site (location_id)
```{r}
race_ratio_locid <- SPL_18 %>%
  group_by(location_id, race) %>%
  dplyr::summarise(count = n()) %>%
  pivot_wider(
    id_cols = location_id,
    names_from = race,
    values_from = count,
    names_prefix = "Observed_"
  ) %>%
  mutate(race_ratio = round(Observed_URM/Observed_White, 2)) %>%
  select(location_id, race_ratio)
```

# Create map by neighborhood - neighborhood fill corresponds to URM/White ratio

## 2018 map
```{r}
# Join the city map with 2018 filtered data ----
race_ratio_18_geo <- left_join(SEA_map, race_ratio, by = "S_HOOD")

ggplot() +
  geom_sf(data = race_ratio_18_geo, aes(fill = race_ratio)) +
  # geom_text(data = FM_ratio_18_geo, aes(label = S_HOOD, x=longitude, y=latitude), size = 3, color = "black", check_overlap = TRUE)+ #Need to troubleshoot
  scale_fill_gradient(name = "URM-to-White staying ratio",
                      low = "lightblue", high = "darkblue", na.value = "gray",
                      guide = guide_colorbar(title.position = "top")) +
  labs(title = "2018 URM-to-White Staying Ratio by Seattle Neighborhood ") +
  theme_minimal()+
  theme(
    plot.title = element_text(family = "Times New Roman", size = 16, face = "bold")
  )

# Save 2018 map as an PNG in the Visualization folder ----
#ggsave(filename = "2018_raceRatio_map.png", path = here("Visualization"), bg="white", height = 7, width = 7)
```

# Create map by dots (dots represent block face; color represents proportion)

## 2018 map
```{r}
# Join location_id coordinates with 2018 filtered data ----
race_ratio_18_locid_geo <- left_join(race_ratio_locid, SPL_1823_geo, by = "location_id")

ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=race_ratio_18_locid_geo, aes(x=longitude, y=latitude, color = race_ratio),shape=20, size=4) + # add points
  #scale_size_continuous(range=c(1,6))+
  scale_color_continuous(name="URM to White Ratio",
                        low = "lightblue", high = "darkblue") +
  labs(title = "2018 URM-to-White Staying Ratio by Observation Site") +
  labs(fill="Observation Sites")

# Save Map_2018_dot in the Visualization folder ----
ggsave(filename = "2018_raceRatio_map_dot.png", path = here("Visualization"), bg="white", width = 7, height = 7)
```


