---
title: "004_SPL_ageprop"
output: html_document
date: "2024-05-20"
---

Create variable for age group proportions, plot dot maps

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
rm(list=ls())
library(tidyverse)
library(psych)
library(dplyr)
library(sf)
library(here)
```

# Import data
```{r}
SPL_1823 <- read.csv(here("data","SPL_1823.csv"))
SPL_1823_geo <- st_read(here("data", "SPL_1823_location.geojson")) # location_id shapefiles and lat/long coordinates
SEA_map <- st_read(here("data/raw","04_SPL_Seattle_Map.geojson")) # base SEA neighborhood map
```

# Recode the age column 
```{r}
# inspect age col
table(SPL_1823$staying_age, SPL_1823$study_id)
# looks like in 2023 they collapsed 25-44 and 45-64 together into 25-64

SPL_1823 <- SPL_1823 %>%
  mutate(
    age=recode(staying_age,
                        "25-44"="25-64",
                        "45-64"="25-64"))

# table(SPL_1823$age)
# table(SPL_1823$staying_age)
```

# Calculate age proportion by observation site (location_id)
```{r}
age_prop_locid <- SPL_1823 %>%
  group_by(location_id, age, study_id) %>%
  dplyr::summarise(count= n()) %>%
  mutate(age = ifelse(age == "", "NA", age)) %>% #replace the empty strings with NA otherwise it won't pivot
  pivot_wider(
    id_cols = c(location_id, study_id),
    names_from = age,
    values_from = count) %>%
  rowwise() %>% #this is to make sure the total is calculated correctly in the next line
  dplyr::mutate(total = sum(`0-4`, `15-24`, `25-64`, `5-14`, `65+`, `NA`, na.rm=T)) %>% #calculate total observed
  mutate_at(vars(`0-4`, `15-24`, `25-64`, `5-14`, `65+`, `NA`), ~replace_na(., 0)) %>% #replace NAs with 0s so that the calculated proportions return 0 and not NA
  mutate(toddler_prop = `0-4`/total,
         tween_prop = `5-14`/total,
         ya_prop = `15-24`/total, #ya stands for young adult
         adult_prop = `25-64`/total,
         senior_prop = `65+`/total) %>%
  select(location_id, study_id, toddler_prop, tween_prop, ya_prop, adult_prop, senior_prop, total)

# append S_HOOD back
age_prop_locid <- SPL_1823 %>%
  select(location_id, S_HOOD) %>%
  distinct() %>%
  right_join(age_prop_locid, by = "location_id")
```

# Create map by dots

## 2018 static map (dots represent each block face; color represent percentage of age group)

```{r}
# Filter 2018 data ----
age_prop_18_locid <- age_prop_locid %>% filter(study_id == "2018_Seattle_Citywide")

# Join location_id coordinates with 2018 filtered data ----
age_prop_18_locid_geo <- left_join(age_prop_18_locid, SPL_1823_geo, by = "location_id")

# Toddlers ----
ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=age_prop_18_locid_geo,
             aes(x=longitude, y=latitude, color = toddler_prop, alpha = 0.95),
             shape=20,
             size=3) + # add points
  scale_color_gradientn(name="%: Toddler/Total",
                        colours = c("slategray1","royalblue4"),
                        breaks = c(0, 0.02, 0.04, 0.06, 0.08, 0.1)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Percentage of Toddlers Staying by Observation Site") +
  labs(fill="Observation Sites")

ggsave(filename = "2018_toddler_prop_map_dot.png", path = here("Visualization"), bg="white", width = 7, height = 7)

# Tween ----
ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=age_prop_18_locid_geo,
             aes(x=longitude, y=latitude, color = tween_prop, alpha = 0.95),
             shape=20,
             size=3) + # add points
  scale_color_gradientn(name="%: Tween/Total",
                        colours = c("slategray1","royalblue4"),
                        breaks = c(0, 0.05, 0.10, 0.15, 0.20)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Percentage of Tweens Staying by Observation Site") +
  labs(fill="Observation Sites")

# Young Adults  ----
ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=age_prop_18_locid_geo,
             aes(x=longitude, y=latitude, color = ya_prop, alpha = 0.95),
             shape=20,
             size=3) + # add points
  scale_color_gradientn(name="%: Young Adults/Total",
                        colours = c("slategray1","royalblue4"),
                        breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Percentage of Young Adults Staying by Observation Site") +
  labs(fill="Observation Sites")

# Adults  ----
ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=age_prop_18_locid_geo,
             aes(x=longitude, y=latitude, color = adult_prop, alpha = 0.95),
             shape=20,
             size=3) + # add points
  scale_color_gradientn(name="%: Adults/Total",
                        colours = c("slategray1","royalblue4"),
                        breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1.0)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Percentage of Adults Staying by Observation Site") +
  labs(fill="Observation Sites")

# Senior  ----
ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=age_prop_18_locid_geo,
             aes(x=longitude, y=latitude, color = senior_prop, alpha = 0.95),
             shape=20,
             size=3) + # add points
  scale_color_gradientn(name="%: Seniors/Total",
                        colours = c("slategray1","royalblue4"),
                        breaks = c(0, 0.10, 0.20, 0.30, 0.40)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Percentage of Seniors Staying by Observation Site") +
  labs(fill="Observation Sites")

# Save Map_2018_dot in the Visualization folder ----
# ggsave(filename = "2018_toddler_prop_map_dot.png", path = here("Visualization"), bg="white", width = 7, height = 7)


```

# 2018 Interactive Plots 
## I am only showing Fem-presenting, but when you hover over, the text box should show M_prop, and also the total number of observations
```{r}
# Fem-presenting
p <- ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=FM_prop_18_locid_geo, aes(x=longitude, y=latitude, color = F_prop, 
text = paste("Block Face: ", location_id, 
             "<br>Percentage of Fem_presenting: ", round(F_prop,2), 
             "<br>Percentage of Masc_presenting: ", round(M_prop,2),
             "<br>Total observation: ", total), 
                                            alpha = 0.95),shape=20, size=4) + # add points
  scale_color_gradientn(name="%: Fem-presenting/Total", colours = c("mistyrose","violetred3"), #Specify the colors for each end 
                       breaks = c(0, 0.2, 0.4, 0.6, 0.8)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Percentage of Feminine-Presenting Individuals Staying by Observation Site") +
  labs(fill="Observation Sites")

p <- ggplotly(, tooltip = "text")

saveWidget(p, here("Interactive", "2018_Gender_interactive.html"), selfcontained = FALSE, libdir = "libs")

```


# 2023 Interactive Plots 
## I am only showing Fem-presenting, but when you hover over, the text box should show M_prop, and also the total number of observations
```{r}
# Fem-presenting
p <- ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=FM_prop_23_locid_geo, aes(x=longitude, y=latitude, color = F_prop, 
text = paste("Block Face: ", location_id, 
             "<br>Percentage of Fem_presenting: ", round(F_prop,2), 
             "<br>Percentage of Masc_presenting: ", round(M_prop,2),
             "<br>Total observation: ", total), 
                                            alpha = 0.95),shape=20, size=4) + # add points
  scale_color_gradientn(name="%: Fem-presenting/Total", colours = c("mistyrose","violetred3"), #Specify the colors for each end 
                       breaks = c(0, 0.2, 0.4, 0.6, 0.8)) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2023 Percentage of Feminine-Presenting Individuals Staying by Observation Site") +
  labs(fill="Observation Sites")

p <- ggplotly(, tooltip = "text")

saveWidget(p, here("Interactive", "2023_Gender_interactive.html"), selfcontained = FALSE, libdir = "libs")

```








