---
title: "005_SPL_FMpropr"
output: html_document
date: "2024-05-17"
---

Recodes gender, calculates proportion of female and male by location_id, and maps the dot map

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load required packages
```{r}
rm(list=ls())
library(tidyverse)
library(psych)
library(dplyr)
library(sf)
library(here)
library(leaflet)
```

# Import data
```{r}
SPL_1823 <- read.csv(here("data","SPL_1823.csv"))
SPL_1823_geo <- st_read(here("data", "SPL_1823_location.geojson")) # location_id shapefiles and lat/long coordinates
SEA_map <- st_read(here("data/raw","04_SPL_Seattle_Map.geojson")) # base SEA neighborhood map
```

# Recode the gender column 
```{r}
SPL_1823 <- SPL_1823 %>%
  mutate(
    gender=recode(staying_gender,
                        "Female"="Fem",
                        "Feminine presenting"="Fem",
                        "Gender-non-conforming presenting" = "non_conforming",
                        "Male"="Masc",
                        "Masculine presenting"="Masc",
                        "Other_Unsure"="Unsure",
                        "Unsure"="Unsure"))

# table(SPL_1823$gender)
# table(SPL_1823$staying_gender)
```

# Calculate proportion women or men by neighborhood (S_HOOD)
```{r}
# instead of doing a f:m ratio, calculate the proportion of total observed that are women/men
FM_prop_shood <- SPL_1823 %>%
  group_by(S_HOOD, gender, study_id) %>%
  dplyr::summarise(count= n()) %>%
  pivot_wider(
    id_cols = c(S_HOOD, study_id),
    names_from = gender,
    values_from = count,
    names_prefix = "Observed_"
  ) %>%
  rowwise() %>% #this is to make sure the total is calculated correctly in the next line
  dplyr::mutate(total = sum(c_across(Observed_Fem:Observed_non_conforming), na.rm=T)) %>% #calculate total observed, including observations where no one was observed!
  mutate_at(vars(Observed_Fem, Observed_Masc, Observed_non_conforming, Observed_Unsure), ~replace_na(., 0)) %>% #replace NAs with 0s so that the calculated proportions return 0 and not NA
  mutate(F_prop = Observed_Fem/total,
         M_prop = Observed_Masc/total) %>%
  select(S_HOOD, study_id, F_prop, M_prop, Observed_Fem, Observed_Masc, total)

FM_prop_shood
# note that the total is the sum of fem, masc, non_conforming, unsure, and empty observations (i.e., empty locations where no one was staying at all). this is why e.g., Bitter Lake in 2018 has 1 fem, 0 masc, and 16 total observations
```

# Calculate proportion women or men by observation site (location_id)

```{r}
# instead of doing a f:m ratio, calculate the proportion of total observed that are women/men
FM_prop_locid <- SPL_1823 %>%
  group_by(location_id, gender, study_id) %>%
  dplyr::summarise(count= n()) %>%
  pivot_wider(
    id_cols = c(location_id, study_id),
    names_from = gender,
    values_from = count,
    names_prefix = "Observed_"
  ) %>%
  rowwise() %>% #this is to make sure the total is calculated correctly in the next line
  dplyr::mutate(total = sum(c_across(Observed_Fem:Observed_non_conforming), na.rm=T)) %>% #calculate total observed
  mutate_at(vars(Observed_Fem, Observed_Masc, Observed_non_conforming, Observed_Unsure), ~replace_na(., 0)) %>% #replace NAs with 0s so that the calculated proportions return 0 and not NA
  mutate(F_prop = Observed_Fem/total,
         M_prop = Observed_Masc/total) %>%
  select(location_id, study_id, F_prop, M_prop, Observed_Fem, Observed_Masc, total)

# append S_HOOD back
FM_prop_locid <- SPL_1823 %>%
  select(location_id, S_HOOD) %>%
  distinct() %>%
  right_join(FM_prop_locid, by = "location_id")

FM_prop_locid
# again, note that the total is the sum of fem, masc, non_conforming, unsure, and empty observations (i.e., empty locations where no one was staying at all). this is why e.g., BLV1 in 2018 has 1 fem, 0 masc, and 8 total observations

```

# Create map by dots - Annie to create template

Right now it's stil the old FMratio map scripts below

## 2018 map
```{r}
# Filter 2018 data ----
FM_ratio_18_locid <- FM_ratio_locid %>% filter(study_id == "2018_Seattle_Citywide")

# Join location_id coordinates with 2018 filtered data ----
FM_ratio_18_locid_geo <- left_join(FM_ratio_18_locid, SPL_1823_geo, by = "location_id")

ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=FM_ratio_18_locid_geo, aes(x=longitude, y=latitude, color = FM_ratio, 
                                             alpha = 0.95),shape=20, size=4) + # add points
  scale_color_gradientn(name="Female to Male Staying Ratio",
                       colours = c("brown","lightyellow","blue"), #Specify the colors for each end 
                       breaks = c(0, 0.5, 1, 1.5, 2, max(FM_ratio_18_locid_geo$FM_ratio))) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2018 Female-to-Male Staying Ratio by Observation Site") +
  labs(fill="Observation Sites")

# Save Map_2018_dot in the Visualization folder ----
ggsave(filename = "2018_FMratio_map_dot.png", path = here("Visualization"), bg="white", width = 7, height = 7)
```

## 2023 map
```{r}
# Filter 2023 data ----
FM_ratio_23_locid <- FM_ratio_locid %>% filter(study_id == "2023_Seattle_Citywide")

# Join location_id coordinates with 2023 filtered data ----
FM_ratio_23_locid_geo <- left_join(FM_ratio_23_locid, SPL_1823_geo, by = "location_id")

ggplot() +
  geom_sf(data = SEA_map) + #plot base SEA neighborhood map
  geom_point(data=FM_ratio_23_locid_geo, aes(x=longitude, y=latitude, color = FM_ratio, alpha = 0.9),shape=20, size=4) + # add points
  scale_color_gradientn(name="Female to Male Staying Ratio",
                       colours = c("brown","lightyellow","blue"), #Specify the colors for each end 
                       breaks = c(0, 0.5, 1, 1.5, 2, max(FM_ratio_23_locid_geo$FM_ratio))) + #Create the breaks
  scale_alpha_continuous(guide = "none")+
  labs(title = "2023 Female-to-Male Staying Ratio by Observation Site") +
  labs(fill="Observation Sites")


# Save Map_2023_dot in the Visualization folder ----
ggsave(filename = "2023_FMratio_map_dot.png", path = here("Visualization"), bg="white", width = 7, height = 7)

```
