---
title: "SPL_demographics"
author: "Rebecca Schachtman"
date: "4/29/2024"
output: html_document
---
# Load libraries
```{r}
library(tidyverse)
library(here)
```

# Load in data files
```{r}
#Census data from: https://data-seattlecitygis.opendata.arcgis.com/datasets/SeattleCityGIS::selected-demographic-and-housing-estimates-dp05/about

#Census data from Seattle from 2010-2022
demo_data <- read.csv(here("data","05_SPL_Demographics.csv")) 

#Clean SPL 2018 and 2023 data
SPL_data <- read.csv(here("data", "SPL_1823.csv")) 
```

# Updating necessary neighborhood names
```{r}
demo_data <- demo_data %>% mutate(neighborhood = recode(Community.Reporting.Area.Name,
                                 "Alki/Admiral" = "Alki",
                                 "Downtown Commercial Core" = "Commercial Core",
                                 "West Seattle Junction/Genesee Hill" = "West Seattle Junction",
                                 "Broadview/Bitter Lake" = "Bitter Lake Village",
                                 "Greenwood/Phinney Ridge" = "Greenwood - Phinney Ridge",
                                 "First Hill" = "First Hill / 12th Ave",
                                 "Highland Park" = "Westwood - Highland Park",
                                 "Roxhill/Westwood" = "Westwood - Highland Park",
                                 "Miller Park" = "Madison Miller",
                                 "Mt. Baker/North Rainier" = "Mt Baker",
                                 "North Beacon Hill/Jefferson Park" = "North Beacon Hill",
                                 "Northgate/Maple Leaf" = "Northgate",
                                 "Olympic Hills/Victory Heights" = "Lake City",
                                 "Wedgwood/View Ridge" = "Wedgwood"))

# Questions:
## Broadview/Bitter Lake, Pioneer Square/ID, and Ravenna/Bryant recode (should I pick one or the other neighborhood to recode it into bc the location_neighborhood column in the SPL_1823 data I'm matching has these as separate entries)
## This dataset only has Queen Anne, but this is split into North and Lower in the location_neighborhood column

## Note that need to update original cleaning code to fix= First Hill / 12th Ave (not broadview) and 23rd and jackson (not pinehurst)
```
**Note:**  
These "Community.Reporting.Area.Names" values from the census data will be dropped in the merge these because there is not a suitable neighborhood to recode them into:  
Arbor Heights, Cascade/Eastlake (closest is SLU), Cedar Park/Meadowbrook, Duwamish/SODO, Central Area/Squire Park, Fauntleroy/Seaview, Green Lake, Haller Ridge, High Point, Interbay, Judkins Park, Laurelhurst/Sand Point, Licton Springs, Madrona/Leschi, Montlake/Portage Bay, North Beach/Blue Ridge, North Capitol Hill, North Delridge, Riverview, South Beacon Hill/NewHolly, Sunset Hills/Loyal Heights

# Create demographic dataframe represents the total counts for census data in each demographic category per neighborhood
```{r}
demo_data_avg <- demo_data %>%
  group_by(neighborhood) %>%
  summarise(across(Total.population:Total.Housing.Units, sum))
```
# Merging demographic data into SPL data
```{r}
# 2018 data
##Creating a 2018 specific staying dataframe
SPL2018 <- SPL_data %>% filter(study_id == "2018_Seattle_Citywide") %>%
  rename("neighborhood" = "location_neighborhood")
##Filtering census data to 2020 only (closest year to 2018 that is available)
demo2018 <- demo_data %>% filter(ACS.Vintage == "5Y20")
##Merging both dataframes by the neighborhood column
merged2018 <- merge(SPL2018, demo2018, by = "neighborhood", all = FALSE)
##Filtering so that there is just one row per unique_staying_id
merged2018 <- distinct(test, unique_staying_id, .keep_all = TRUE)

# 2023 data
##Creating a 2023 specific staying dataframe
SPL2023 <- SPL_data %>% filter(study_id == "2023_Seattle_Citywide") %>%
  rename("neighborhood" = "location_neighborhood")
##Filtering census data to 2022 only (closest year to 2023 that is available)
demo2023 <- demo_data %>% filter(ACS.Vintage == "5Y22")
##Merging both dataframes by the neighborhood column
merged2023 <- merge(SPL2018, demo2018, by = "neighborhood", all = FALSE)
##Filtering so that there is just one row per unique_staying_id
merged2023 <- distinct(test, unique_staying_id, .keep_all = TRUE)

# Creating one big dataframe for 2018 and 2023 data
SPL_1823_demographics <- rbind(merged2018, merged2023)
##Cleaning it up
SPL_1823_demographics <- SPL_1823_demographics %>% dplyr::select(neighborhood, unique_staying_id:ACS.Vintage)
##Saving it
write.csv(SPL_1823_demographics, here("data", "SPL_1823_demographics.csv"))
```

